---
title: "Flume: Random Analysis (March-November)"
subtitle: "(adapted from Jarad Niemi - Soilpad Analysis)"
author: "Jessica Nelson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("lme4")
library("lmerTest")
library("tidyverse"); theme_set(theme_bw())
library("lattice")
library("emmeans")
library("ggResidpanel")
library("data.table")
library("stringr")
library("ggplot2")
library("ggpmisc")
library("performance")
library("multcomp")
library("xtable")


options(width = 120, scipen = 999)
windowsFonts(A = windowsFont("Times New Roman")) 

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```

## Read in data

```{r, child="sediment.Rmd"}
```

```{r load, dependson= "full_dataframe"}
load("full_df.RData")
```

```{r flume_random, dependson = "load"}
flumeR <- full_df %>%
  #filter(!is.na(ro_event)) %>%
    subset(random == 'R') %>%
  mutate(Treatment = factor(Treatment, levels=c('strips','control')))

```

## Exploratory analysis

## Site-year with rainfall event

```{r, dependson="load"}
site_year_rfevent <- flumeR %>%
  dplyr::select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_rfevent, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```


```{r, dependson="flume_random"}
site_year_rfeventR <- flumeR %>%
  dplyr::select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_rfeventR, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```


## Number of samples

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r sample_event_counts, dependson="flume_random"}
TSS_countsR <- flumeR %>%
  group_by(Year, SiteID, sample_event) %>%
  distinct() %>%
  summarize(n = n(), .groups = "drop")
```


```{r year_tss_sum, dependson="flume_random"}
TSS_sumR <- flumeR %>%
  group_by(Year, SiteID, Treatment) %>%
  distinct() %>%
  summarize(tss_sum = sum(tss_sum), .groups = "drop")
```

Plot the number of observations for each combination. 

```{r plot_count, dependson="sample_event_counts", out.width = "\\textwidth"}
g <- ggplot(TSS_countsR, aes(x = Year, y= SiteID, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  scale_fill_gradient(low = "blue", high = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g 
##ggsave("fig/soilpad_counts_no_diversion.png", g, width = 12, height = 12)
```


# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss


```{r designR_model, dependson = "flume_random"}
mR_flume <- lmerTest::lmer(log(tss_sum+0.001) ~
                             Treatment*ln_ppt +
                            Treatment*crop +
                            Year*Treatment +
                            #(1 | SiteID) + #removed due to singular fit
                             (1 | SiteID:Treatment) + #removed due to singular fit
                            #(1|Year:sample_event) + #removed due to singular fit
                            (1|SiteID:Year:sample_event),
                           
                          data = flumeR)


summary(mR_flume)
anova(mR_flume)
performance::r2(mR_flume)
#confint(mR_flume, oldNames = FALSE)
r2_nakagawa(mR_flume)
icc(mR_flume)

```

```{r designR_step_model, dependson = "designR_model"}
mR_flume_step <- step(mR_flume, reduce.random = FALSE, alpha.fixed = 0.1)
mR_flume_model <- get_model(mR_flume_step)
summary(mR_flume_model)
anova(mR_flume_model)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
```


```{r designR_model_contrasts, dependson = "designR_model"}
trt_yrR = emmeans(mR_flume, pairwise ~ Treatment|Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(trt_yrR)$contrasts
trt_yrR$contrasts

trtR    = emmeans(mR_flume, pairwise ~ Treatment, 
                    type = "response", 
                    lmer.df = "asymptotic")
confint(trtR)

yearR   = emmeans(mR_flume, ~ Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(yearR)

trt_pptR  = emmeans(mR_flume, pairwise ~ Treatment|ln_ppt,
                    at=list(ln_ppt=c(-4,-3,-2,-1,-0.25)), 
                    type = "response",
                    lmer.df = "asymptotic")

confint(trt_pptR)$contrasts ## exp. the values

cropR      = emmeans(mR_flume, pairwise ~ Treatment|crop,   
                    type = "response",
                    lmer.df = "asymptotic")

confint(cropR)$contrasts
```


```{r, dependson="designR_model_contrasts"}
cropR_em <- as.data.frame(cropR)
 
kR <- cropR_em %>%
  filter(contrast != "strips - control")
 
cropR_em_plot <- kR %>%
   ggplot(aes(x=Treatment, y=response, fill=crop))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (response-SE), ymax = (response+SE)),
             width = 0.2,
             linetype = "solid",
             position = position_dodge(width = 0.5),
             color="black", size=0.7) +
   scale_fill_manual(values = c("corn" = "#FFCC33",
                              "soybean" = "#339900"), 
                    labels = c("Corn", "Soybean")) +
   ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset)") +
   xlab("Treatment") +
   ylab("TSS Load \n(kg/ha)") +
   theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
         axis.title.x = element_text(size=18, face="bold"),
         axis.title.y = element_text(size=18, face="bold"),
         axis.text.x = element_text(size=18),
         axis.text.y = element_text(size=18)) #+
  #scale_x_discrete(labels= c("Control", "Prairie Strip")) #+
  #scale_y_continuous(limits = c(0, 40))
 
cropR_em_plot

```



```{r, dependson="load"}
h <- ggplot(flumeR, aes(x=log(rain), y=log(tss_sum+0.001)), inherit.aes = FALSE) +
#h <- ggplot(flumeR, aes(x=LSfactor, y=log(tss_sum+0.001)), inherit.aes = FALSE) +
 scale_color_manual(values = c("control" = "#C38820",
                              "strips" = "#176D9C"),
                    labels=c("Control", "Prairie Strip"),
                    name  ="") +
 geom_point(aes(color=Treatment), size=6) +
 xlab("log(Rainfall Total (mm))") +
 ylab("log(TSS Load (kg/ha))") +
 geom_smooth(method=lm, se=FALSE, fullrange=TRUE, size=2, color="black") +
 ggtitle("B. Randomized Location Subset") +
 theme(plot.title = element_text(family="serif", size=28, face="bold", hjust=0.5),
       axis.title.x = element_text(family="serif", size=28, face="bold"),
       axis.title.y = element_text(family="serif", size=28, face="bold"),
       axis.text.x = element_text(family="serif", size=28),
       axis.text.y = element_text(family="serif", size=28),
       legend.text = element_text(family = "serif", size=28),
       legend.key.size = unit(2,"line")) +
 stat_poly_eq(formula = y ~ x,
              aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")),
              parse = TRUE,
              label.x.npc = "left",
              label.y.npc = "top",
              vstep = 0.05,
              size = 8)+
  scale_y_continuous(limits = c(-7,8)) +
  scale_x_continuous(limits = c(-5,-1))


h
ggsave("../code/fig/randReg_ppt_load.png", h, width = 12, height = 12)

```

# Check assumptions

There are two possible models:

- mR_flume: full model design, design-based analysis
- mR_flume_model: model design selected based on backward step selection

### Full model design

```{r designR_model_residual_plots, dependson = "designR_model"}
resid_panel(mR_flume, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mR_flume)
```

```{r designR_step_residual_plots, dependson = "designR_model"}
resid_panel(mR_flume_model, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mR_flume_model)
```

```{r designR_model_figure, dependson="load"}
groupR <- flumeR %>% 
  group_by(Treatment) %>% 
  summarize(n = n(), 
            mean = mean(tss_sum), 
            sd = sd(tss_sum),
            .groups = "drop") %>%
  mutate(se = sd / sqrt(n),
         lb = mean + qt(0.025, df = n-1)*se+1.7,
         ub = mean - qt(0.025, df = n-1)*se)

groupR[groupR < 0] <- 0

trt_plotR <- groupR %>%
  ggplot(aes(x=Treatment, y=mean, fill=Treatment))+
  geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
  geom_errorbar(aes(ymin = (lb), ymax = (ub)),
                width = 0.3,
                linetype = "solid",
                position = position_dodge(width = 0.5),
                color="black", size=0.7) +
  scale_fill_manual(values = c("control" = "#C38820",
                              "strips" = "#176D9C")) +
  ggtitle("B. Randomized Location Subset") +
  xlab("Treatment") +
  ylab("TSS Load (kg/ha)") +
  theme(plot.title = element_text(family = "serif", size=28, face="bold", hjust=0.5),
       axis.title.x = element_text(family = "serif", size=28, face="bold"),
       axis.title.y = element_text(family = "serif", size=28, face="bold"),
       axis.text.x = element_text(family = "serif", size=28),
       axis.text.y = element_text(family = "serif", size=28)) +
  scale_x_discrete(labels= c("Prairie Strip", "Control")) +
  scale_y_continuous(limits = c(0, 50)) +
  theme(text = element_text())

trt_plotR
ggsave("../code/fig/randflume_plot.png", trt_plotR)
```

```{r mean_tss_figure, dependson="load"}
crop_em <- as.data.frame(crop)

k <- crop_em %>%
  filter(contrast != "strips - control")

crop_em_plot <- k %>%
   ggplot(aes(x=crop, y=response, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (response-SE), ymax = (response+SE)),
             width = 0.2,
             linetype = "solid",
             position = position_dodge(width = 0.5),
             color="black", size=0.7) +
   #scale_fill_manual(values = c("corn" = "#FFCC33",
  #                            "soybean" = "#339900"),
  #                  labels = c("Corn", "Soybean")) +
   ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset)") +
   xlab("Treatment") +
   ylab("TSS Load \n(kg/ha)") +
   theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
         axis.title.x = element_text(size=18, face="bold"),
         axis.title.y = element_text(size=18, face="bold"),
         axis.text.x = element_text(size=18),
         axis.text.y = element_text(size=18)) #+
  #scale_x_discrete(labels= c("Control", "Prairie Strip")) #+
  #scale_y_continuous(limits = c(0, 40))

crop_em_plot


trtRcrp_grp <- flumeR %>% 
  group_by(crop, Treatment) %>% 
   summarize(n = n(), 
             mean = mean(tss_sum), 
             sd = sd(tss_sum),
             groups = "drop") %>%
   mutate(se = sd / sqrt(n),
          lb = mean + qt(0.025, df = n-1)*se,
          ub = mean - qt(0.025, df = n-1)*se)
 
trtRcrp_grp[trtRcrp_grp < 0] <- 0

 trtRcrp_plot <- trtRcrp_grp %>%
   ggplot(aes(x=crop, y=mean, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (mean-se), ymax = (mean+se)),
                 width = 0.3,
                 linetype = "solid",
                 position = position_dodge(width = 0.5),
                 color="black", size=0.7) +
   scale_fill_manual(values = c("control" = "#C38820",
                               "strips" = "#176D9C")) +
   ggtitle("B. Randomized Location Subset") +
   xlab("Crop Planted") +
   ylab("Mean TSS Load (kg/ha)") +
  theme(plot.title = element_text(family="serif", size=28, face="bold", hjust=0.5),
        axis.title.x = element_text(family="serif", size=28, face="bold"),
        axis.title.y = element_text(family="serif", size=28, face="bold"),
        axis.text.x = element_text(family="serif", size=28),
        axis.text.y = element_text(family="serif", size=28)) +
   scale_x_discrete(labels= c("Corn", "Soybean")) +
   scale_y_continuous(limits = c(0, 70))
 
trtRcrp_plot
ggsave("../code/fig/randCrop_load_plot.png", trtRcrp_plot)

```



``` {r mean_tss_figure, dependson="load"}
trtyrR_grp <- flumeR %>% 
  group_by(Year, Treatment) %>% 
   summarize(n = n(), 
             mean = mean(tss_sum), 
             sd = sd(tss_sum),
             groups = "drop") %>%
   mutate(se = sd / sqrt(n),
          lb = mean + qt(0.025, df = n-1)*se,
          ub = mean - qt(0.025, df = n-1)*se)

str2017 <- data.frame(factor(2017), "strips", 0, 0, 0, "drop", 0, 0, 0)
names(str2017) <- c("Year", "Treatment", "n", "mean", "sd", "groups", "se", "lb", "ub")
trtyrR_grpNEW <- rbind(trtyrR_grp, str2017)

ctrl2017 <- data.frame(factor(2017), "control", 0, 0, 0, "drop", 0, 0, 0)
names(ctrl2017) <- c("Year", "Treatment", "n", "mean", "sd", "groups", "se", "lb", "ub")
trtyrR_grpNEW <- rbind(trtyrR_grpNEW, ctrl2017)

trtyrR_grpNEW[trtyrR_grpNEW==0] <- NA
trtyrR_grpNEW[trtyrR_grpNEW < 0] <- 0

trtR_plot <- trtyrR_grpNEW %>%
  ggplot(aes(x=Year, y=mean, fill=Treatment))+
  geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
  geom_errorbar(aes(ymin = (mean-se), ymax = (mean+se)),
                width = 0.3,
                linetype = "solid",
                position = position_dodge(width = 0.5),
                color="black", size=0.7) +
  scale_fill_manual(values = c("control" = "#C38820",
                             "strips" = "#176D9C")) +
  ggtitle("B. Randomized Location Subset") +
  xlab("Treatment") +
  ylab("TSS Load (kg/ha)") +
  theme(plot.title = element_text(family="serif", size=28, face="bold", hjust=0.5),
        axis.title.x = element_text(family="serif", size=28, face="bold"),
        axis.title.y = element_text(family="serif", size=28, face="bold"),
        axis.text.x = element_text(family="serif", size=20),
        axis.text.y = element_text(family="serif", size=20)) +
   #scale_x_discrete(labels= c("Prairie Strip", "Control"))+
   scale_y_continuous(limits = c(0, 125))
 
 trtR_plot

 ggsave("../code/fig/randflumeyr_plot.png", trtR_plot)
 
```


```{r extract-contrast-ci, dependson="design_model"}
extract_contrast_ci <- function(d) {
  mR <- lmerTest::lmer(update(mR_flume, 
                          as.formula(paste0("~.+",d$variable))),
                    data = flumeR) 
    
  # Obtain coefficient for new variable
  av <- anova(mR)
  rowid <- which(rownames(av) == d$variable)
  
   
  em <- emmeans(mR, 
                pairwise ~ Treatment, 
                type = "response", 
                lmer.df = "asymptotic")
  
  cbind(
    av[rowid,] %>% as.data.frame,
    confint(em)$contrasts %>% as.data.frame
  )
}
```


```{r treatment-effect-with-explanatory-variables, dependson=c("extract-contrast-ci")}
add <- data.frame(variable = c("ln_slp", "C_factor", "K_factor", "LSfactor", "Lfactor", "Sfactor", "ln_S", "ln_L", "ln_LS", "slope75", "ln_postL", "ln_postT"))


added_variable_analysesR <- add %>% 
  group_by(variable) %>%
  do(extract_contrast_ci(.))
```


```{r print-added-variable-analysis, dependson="treatment-effect-with-explanatory-variables", results='asis'}
added_variable_analysesR %>%
  xtable() %>%
  print(type="html", include.rownames=FALSE)
```