---
title: "Flume: Full Analysis (March-November)"
subtitle: "(adapted from Jarad Niemi - Soilpad Analysis)"
author: "Jessica Nelson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("lme4")
library("lmerTest")
library("tidyverse"); theme_set(theme_bw())
library("lattice")
library("emmeans")
library("ggResidpanel")
library("data.table")
library("stringr")
library("ggplot2")
library("ggpmisc")
library("performance")
library("multcomp")
library("xtable")


options(width = 120, scipen = 999)
windowsFonts(A = windowsFont("Times New Roman")) 

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```

## Read in data
```{r load, dependson= "full_dataframe"}
full_out <- read_csv("../data/tidy/full_df_out.csv") %>%
  mutate(Year = factor(Year),
         Treatment = factor(Treatment, levels=c('strips','control')))

full_in <- read_csv("../data/tidy/full_df_in.csv") %>%
  mutate(Year = factor(Year),
         Treatment = factor(Treatment, levels=c('strips','control')))

```


## Exploratory analysis

## Flume Out: Site-year with sample event

```{r, dependson="load"}
site_year_eventOut <- full_out %>%
  dplyr::select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_eventOut, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```

## Flume In: Site-year with sample event

```{r, dependson="load"}
site_year_eventIn <- full_in %>%
  dplyr::select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_eventIn, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```

## Data visualization

## Number of samples: Flume Out

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r sample_event_counts, dependson="load"}
out_counts <- full_out %>%
  group_by(Year, SiteID, Treatment, random) %>%
  distinct() %>%
  summarize(n = n(), .groups = "drop")
```

Plot the number of observations for each combination. 

```{r plot_count, dependson="sample_event_counts", out.width = "\\textwidth"}
gOut = ggplot(out_counts, aes(x = Treatment, y = SiteID, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  facet_grid(random ~ Year) +
  scale_fill_gradient(low = "blue", high = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

gOut

```


```{r mean_data, dependson = "load"}
load_out <- full_out %>%
  group_by(Year, SiteID, Treatment, crop) %>%
  summarize(sum_load             = sum(tss_load, na.rm = TRUE),
            log_load = mean(log(sum_load), na.rm = TRUE),
            n                   = n(),
            .groups = "drop")

```


```{r wp_per_day_plot, dependson = "plot_data"}
gOut <- ggplot(load_out, 
           aes(x = Treatment, 
               y = sum_load)) + 
  geom_point() + 
  geom_line() +
  facet_grid(SiteID ~ Year, scales = "free_y") + 
  labs(title ="Total load (kg/ha) per field season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
gOut

#ggsave("fig/wp_per_day_plot.png", gOut)
```


## Number of samples: Flume In

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r sample_event_counts, dependson="load"}
in_counts <- full_in %>%
  group_by(Year, SiteID, Treatment, random) %>%
  distinct() %>%
  summarize(n = n(), .groups = "drop")
```

Plot the number of observations for each combination. 

```{r plot_count, dependson="sample_event_counts", out.width = "\\textwidth"}
gIn = ggplot(in_counts, aes(x = Treatment, y = SiteID, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  facet_grid(random ~ Year) +
  scale_fill_gradient(low = "blue", high = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

gIn
```


```{r mean_data, dependson = "load"}
load_in <- full_out %>%
  group_by(Year, SiteID, Treatment, crop) %>%
  summarize(sum_load             = sum(tss_load, na.rm = TRUE),
            log_load = mean(log(sum_load), na.rm = TRUE),
            n                   = n(),
            .groups = "drop")

```


```{r wp_per_day_plot, dependson = "plot_data"}
gIn <- ggplot(load_in, 
           aes(x = Treatment, 
               y = sum_load)) + 
  geom_point() + 
  geom_line() +
  facet_grid(SiteID ~ Year, scales = "free_y") + 
  labs(title ="Total load (kg/ha) per field season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
gIn

#ggsave("fig/wp_per_day_plot.png", gIn)
```


# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss

## Confirmatory, design-based analysis

### Treatment effect: Flume Out

```{r design_model_Out, dependson = "load"}
mOut <- lmerTest::lmer(log(tss_sum+0.001) ~ 
                             Treatment*ln_ppt +
                            Treatment*crop +
                            Year*Treatment +
                            
                            #(1 | SiteID) + #removed due to singular fit
                             (1 | SiteID:Treatment) + 
                            (1|Year:sample_event) + #consider this and below with SiteID 
                            (1|SiteID:Year:sample_event), 
                          
                          data = full_in)

summary(mOut)
confint(mOut, oldNames = FALSE)
anova(mOut)
r2_nakagawa(mOut)
icc(mOut)

```

```{r design_model_Outstep, dependson = "design_model_Out"}
mOut_step <- step(mOut, reduce.random = FALSE, alpha.fixed = 0.1)
mOut_model <- get_model(mOut_step)
summary(mOut_model)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
```

# Check assumptions

There are two possible models:

- m_flume: full model design, design-based analysis
- m_flume_model: model design selected based on backward step selection

### Full model design

```{r design_Outmodel_residual_plots, dependson = "design_model_Out"}
resid_panel(mOut, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mOut)
```

### Selected model design
```{r design_Outstep_residual_plots, dependson = "design_model_Outstep"}
resid_panel(mOut_model, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mOut_model)
```


```{r design_model_Out_contrasts, dependson = "design_model_Out"}
trt_yrO = emmeans(mOut, pairwise ~ Treatment|Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(trt_yrO)$contrasts

trtO = emmeans(mOut, pairwise ~ Treatment, 
                    type = "response", 
                    lmer.df = "asymptotic")
confint(trtO)

yearO = emmeans(mOut, ~ Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(yearO)

cropO      = emmeans(mOut, pairwise ~ Treatment|crop,   
                    type = "response",
                    lmer.df = "asymptotic")

confint(cropO)$contrasts

```


### Treatment effect: Flume In

```{r design_model_In, dependson = "load"}
mIn <- lmerTest::lmer(log(tss_load+0.025) ~ 
                             Treatment*ln_ppt +
                            Treatment*crop +
                            Year*Treatment +
                            
                            #(1 | SiteID) + #removed due to singular fit
                             (1 | SiteID:Treatment) + 
                            (1|Year:sample_event) + #consider this and below with SiteID 
                            (1|SiteID:Year:sample_event), 
                          
                          data = full_in)

summary(mIn)
confint(mIn, oldNames = FALSE)
anova(mIn)
r2_nakagawa(mIn)
icc(mIn)

```


```{r design_modelstep_In, dependson = "design_model_In"}
mIn_step <- step(mIn, reduce.random = FALSE, alpha.fixed = 0.1)
mIn_model <- get_model(mIn_step)
summary(mIn_model)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
```


# Check assumptions

There are two possible models:

- m_flume: full model design, design-based analysis
- m_flume_model: model design selected based on backward step selection

### Full model design

```{r design_Inmodel_residual_plots, dependson = "design_model_In"}
resid_panel(mOut, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mOut)
```

### Selected model design
```{r design_Instep_residual_plots, dependson = "design_modelstep_In"}
resid_panel(m_flume_model, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(m_flume_model)
```


```{r design_model_In_contrasts, dependson = "design_model_In"}
trt_yrI = emmeans(mIn, pairwise ~ Treatment|Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(trt_yrI)$contrasts

trtI = emmeans(mIn, pairwise ~ Treatment, 
                    type = "response", 
                    lmer.df = "asymptotic")
confint(trtI)

yearI = emmeans(mIn, ~ Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(yearI)

cropI      = emmeans(mIn, pairwise ~ Treatment|crop,   
                    type = "response",
                    lmer.df = "asymptotic")

confint(cropI)$contrasts

```


## Analysis results visualization

## Mean TSS load: Flume Out

```{r mean_tss_figure_Out, dependson="design_model_Out_contrasts"}
trt_frameO <- as.data.frame(trtO)
 
kO <- trt_frameO %>%
  filter(contrast != "strips - control")
 
 trt_plotO <- kO %>%
   ggplot(aes(x=Treatment, y=response, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (response-asymp.LCL), ymax = (response+asymp.UCL)),
             width = 0.2,
             linetype = "solid",
             position = position_dodge(width = 0.5),
             color="black", size=0.7) +
   scale_fill_manual(values = c("control" = "#176D9C",
                                "strips" = "#C38820")) +
   ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset: Out)") +
   xlab("Treatment") +
   ylab("TSS Load \n(kg/ha)") +
   theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
         axis.title.x = element_text(size=18, face="bold"),
         axis.title.y = element_text(size=18, face="bold"),
         axis.text.x = element_text(size=18),
         axis.text.y = element_text(size=18)) +
   scale_x_discrete(labels= c("Control", "Prairie Strip"))
 
trt_plotO

```

## Mean TSS load: Flume In

```{r mean_tss_figure_In, dependson="design_model_In_contrasts"}
trt_frameI <- as.data.frame(trtI)
 
kI <- trt_frameI %>%
  filter(contrast != "strips - control")
 
 trt_plotI <- kI %>%
   ggplot(aes(x=Treatment, y=response, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (response-asymp.LCL), ymax = (response+asymp.UCL)),
             width = 0.2,
             linetype = "solid",
             position = position_dodge(width = 0.5),
             color="black", size=0.7) +
   scale_fill_manual(values = c("control" = "#176D9C",
                                "strips" = "#C38820")) +
   ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset: In)") +
   xlab("Treatment") +
   ylab("TSS Load \n(kg/ha)") +
   theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
         axis.title.x = element_text(size=18, face="bold"),
         axis.title.y = element_text(size=18, face="bold"),
         axis.text.x = element_text(size=18),
         axis.text.y = element_text(size=18)) +
   scale_x_discrete(labels= c("Control", "Prairie Strip"))
 
trt_plotI

```




```{r extract-contrast-ci, dependson="design_model"}
extract_contrast_ci <- function(d) {
  m <- lmerTest::lmer(update(m_flume, 
                          as.formula(paste0("~.+",d$variable))),
                    data = full_df) 
    
  # Obtain coefficient for new variable
  av <- anova(m)
  rowid <- which(rownames(av) == d$variable)
  
   
  em <- emmeans(m, 
                pairwise ~ Treatment, 
                type = "response", 
                lmer.df = "asymptotic")
  
  cbind(
    av[rowid,] %>% as.data.frame,
    confint(em)$contrasts %>% as.data.frame
  )
}
```


```{r treatment-effect-with-explanatory-variables, dependson=c("extract-contrast-ci")}
add <- data.frame(variable = c("slope75", "C_factor", "K_factor", "LSfactor", "Lfactor", "Sfactor", 
               "post_Lupslp", "post_Tupslp", "ln_ppt", "pre_Lupslp"))

added_variable_analyses <- add %>% 
  group_by(variable) %>%
  do(extract_contrast_ci(.))
```


```{r print-added-variable-analysis, dependson="treatment-effect-with-explanatory-variables", results='asis'}
added_variable_analyses %>%
  xtable() %>%
  print(type="html", include.rownames=FALSE)
```

### Rainfall vs. Load: Flume Out
```{r, dependson="load"}
hOut <- ggplot(full_out, aes(x=log(ppt_sum), y=log(tss_load+0.0016), color=Treatment), inherit.aes = FALSE) +
 scale_color_manual(values = c("control" = "#C38820",
                              "strips" = "#176D9C"),
                  #labels=c( "Control", "Prairie Strip"),
                    name  ="") +
 geom_point(size=5) +
 xlab("log(Rainfall Total (mm))") +
 ylab("log(TSS load (kg/ha))") +
 geom_smooth(method=lm, se=FALSE, fullrange=TRUE, size=3) +
 ggtitle("A. Full Dataset (Out)") +
 theme(plot.title = element_text(family="A", size=28, face="bold", hjust=0.5),
       axis.title.x = element_text(family="A", size=28, face="bold"),
       axis.title.y = element_text(family="A", size=28, face="bold"),
       axis.text.x = element_text(family="A", size=28),
       axis.text.y = element_text(family="A", size=28),
       legend.text = element_text(family = "A", size=28),
       legend.key.size = unit(3,"line")) +
 stat_poly_eq(formula = y ~ x,
              aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")),
              parse = TRUE,
              label.x.npc = "left",
              label.y.npc = "top",
              vstep = 0.05,
              size = 8)

hOut

#ggsave("E:/ISU/Project/SoilMove/data/statistics/flume_analysis/code/fig/Reg_ppt_loadOUT.png", hOut, width = 12, height = 12)

```
