---
title: "Flume: Full Analysis (March-November)"
subtitle: "(adapted from Jarad Niemi - Soilpad Analysis)"
author: "Jessica Nelson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("lme4")
library("lmerTest")
library("tidyverse"); theme_set(theme_bw())
library("lattice")
library("emmeans")
library("ggResidpanel")
library("data.table")
library("stringr")
library("ggplot2")
library("ggpmisc")
library("performance")
library("multcomp")
library("xtable")


options(width = 120, scipen = 999)
windowsFonts(A = windowsFont("Times New Roman"))

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```

## Read in data

```{r, child="sediment.Rmd"}
```


```{r load, dependson= "full_dataframe"}
load("full_df.RData")

#full_df <- read_csv("../data/tidy/full_df_in.csv") %>%
#  mutate(Year = factor(Year),
#         Treatment = factor(Treatment, levels=c('strips','control')))

#load("full_df.RData")
#full_df <- read_csv("../data/tidy/full_df_out.csv") %>%
#  mutate(Year = factor(Year),
#        Treatment = factor(Treatment, levels=c('strips','control')))
  
#load_weight_time$ln_ppt <- log((load_weight_time$precipitation+0.0004)*100)
#load_weight_time$Year <- factor(load_weight_time$Year)

##Statistical significance between treatments
rusle <- read_csv("../data/scratch/rusle_calculations.csv") %>%
    filter(Scenario == "pre-prairie") #%>%
  #filter(random == "R")

control <- rusle$control
strips <- rusle$strips

rusle_st <- rusle %>%
  filter(Scenario == "current") %>%
  group_by(Treatment) %>%
  summarise(
    count = n(),
    mean = mean(A, na.rm = TRUE),
    sd = sd(A, na.rm = TRUE)
  )

res <- t.test(control, strips, paired=TRUE)
res

```

## Exploratory analysis

## Site-year with sample event

```{r, dependson="load"}
site_year_rfevent <- full_df %>%
  dplyr::select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_rfevent, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```

## Data visualization

## Number of samples

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r sample_event_counts, dependson="load"}
TSS_Lfactor <- full_df %>%
  group_by(Lfactor, Sfactor, LSfactor, SiteID, Treatment) %>%
  summarize(ln_mean = mean(log(tss_sum+0.001)),
            ln_sd = sd(log(tss_sum+0.001)),
            mean = mean(tss_sum),
            sd = sd(tss_sum),
            n = n(), 
            .groups = "drop") %>%
  mutate(se = sd / sqrt(n),
         lb = mean + qt(0.025, df = n-1)*se,
         ub = mean - qt(0.025, df = n-1)*se)

slope_factors <- TSS_Lfactor %>%
  gather(value, key = "factor", Lfactor:LSfactor) #%>%
  #spread(key=Treatment, value=c("mean"))

```

Plot the number of observations for each combination. 

```{r plot_count, dependson="sample_event_counts", out.width = "\\textwidth"}
g = ggplot(TSS_counts, aes(x = Treatment, y = crop, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  facet_grid(SiteID + random ~ Year) +
  scale_fill_gradient(low = "blue", high = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g 
```


```{r, dependson="load"}
#h <- ggplot(full_df, aes(x=log(rain), y=log(tss_sum+0.001), color=Treatment), inherit.aes = FALSE) +
#h <- ggplot(full_df, aes(x=Lfactor, y=log(tss_sum+0.001), color=Treatment), inherit.aes = FALSE) +
h <- ggplot(TSS_Lfactor, aes(x=Lfactor, y=mean, color=Treatment), inherit.aes = FALSE) +
 scale_color_manual(values = c("control" = "#C38820",
                              "strips" = "#176D9C"),
                  labels=c( "Control", "Prairie Strip"),
                    name  ="") +
  geom_errorbar(aes(ymin=mean-se, ymax=mean+se), colour="gray", width=.1) +
 geom_point(aes(color=Treatment), size=6) +
  
 xlab("L Factor") +
 ylab("Mean TSS load (log(kg/ha))") +
 geom_smooth(method=lm, se=FALSE, fullrange=TRUE, size=2) +
 ggtitle("A. Full Dataset") +
 theme(plot.title = element_text(family="serif", size=28, face="bold", hjust=0.5),
       axis.title.x = element_text(family="serif", size=28, face="bold"),
       axis.title.y = element_text(family="serif", size=28, face="bold"),
       axis.text.x = element_text(family="serif", size=28),
       axis.text.y = element_text(family="serif", size=28),
       legend.text = element_text(family = "serif", size=28),
       legend.key.size = unit(1,"line")) +
 stat_poly_eq(formula = y ~ x,
              aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")),
              parse = TRUE,
              label.x.npc = "left",
              label.y.npc = "bottom",
              vstep = 0.05,
              size = 8)

h

#ggsave("E:/ISU/Project/SoilMove/data/statistics/flume_analysis/code/fig/Reg_lfactor_load.png", h, width = 12, height = 12)

```

```{r, dependson="load"}
g <- slope_factors %>%
   filter(Treatment == "control") %>%
  ggplot(aes(SiteID, mean))  + 
  geom_bar(aes(x=SiteID, y=mean), stat="identity", fill= "gray") +
  geom_line(aes(x=SiteID, y=value*100, color = factor, group=factor),stat="identity", size=2)+
  labs(title= "Full Dataset: Control Subcatchments",
       x="Site ID",y="Mean TSS Load") +
  scale_y_continuous(sec.axis=sec_axis(~.*0.01,name="Slope Factor"))
g
 
ggsave("../code/fig/control_factors_load.png", g, width = 12, height = 12)

```


```{r, dependson="load"}
gg <- slope_factors %>%
   filter(Treatment == "strips") %>%
  ggplot(aes(SiteID, mean))  + 
  geom_bar(aes(x=SiteID, y=mean), stat="identity", fill= "gray") +
  geom_line(aes(x=SiteID, y=value*100, color = factor, group=factor),stat="identity", size=2)+
  labs(title= "Full Dataset: Prairie Strip Subcatchments",
       x="Site ID",y="Mean TSS Load") +
  scale_y_continuous(sec.axis=sec_axis(~.*0.01,name="Slope Factor"))
gg
 
ggsave("../code/fig/strip_factors_load.png", gg, width = 12, height = 12)

```


```{r mean_data, dependson = "load"}
load_time <- read_csv("../data/tidy/load_weight_year220622.csv")
load_time

model <- lm(log(tss_sum+0.001) ~ Treatment*Lfact_below,
            data = full_df)

summary(model)
anova(model)
```

```{r wp_plot, dependson = "plot_data", eval = FALSE, echo = FALSE}
g <- ggplot(full_df, 
           aes(x = log(rain), y = log(tss_sum+0.001), group=Treatment, color=Treatment)) + 
  geom_line() +
  geom_point() +
  facet_grid(SiteID + random ~ Year) + 
  #scale_y_log10() + 
  labs(title ="Total load (log(kg/ha)) by rainfall accumulation per sample event (log(m))") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_plot.png", g)
```



```{r wp_per_day_plot, dependson = "plot_data"}
g <- ggplot(load_sum, 
           aes(x = Treatment, 
               y = sum_load)) + 
  geom_point() + 
  geom_line() +
  facet_grid(SiteID ~ Year, scales = "free_y") + 
  #scale_y_log10() + 
  labs(title ="Total load (kg/ha) per field season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_per_day_plot.png", g)
```

Average isn't realistic


```{r strip_boxplot, dependson = "create_soilpad", eval = FALSE, echo = FALSE}
g <- ggplot(flume, 
           aes(x = Year, y = rain, color=crop)) + 
  geom_boxplot() +
  geom_point() + 
  #geom_jitter() +
  facet_grid( ~ SiteID) + 
  #scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
g
```

#```{r, dependson="create_sediment"}
#pivot_sample %>% 
#  anova_test(ln_trt ~ ln_ppt*ln_ctl)
## purr https://stackoverflow.com/questions/50702152/compare-models-via-anova-with-purrr-or-dplyr
## anova() and may need an linear model built up.
#```


# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss

## Confirmatory, design-based analysis

### Treatment effect

```{r design_model, dependson = "load"}

m_flume <- lmerTest::lmer(log(tss_sum+0.001) ~ 
                             Treatment*ln_ppt +
                            Treatment*crop +
                            Year*Treatment +
                            #(1 | SiteID) + #removed due to singular fit
                             (1 | SiteID:Treatment) + 
                            (1|Year:sample_event) + #consider this and below with SiteID 
                            (1|SiteID:Year:sample_event), 
                          
                          data = full_df)

summary(m_flume)
anova(m_flume)
performance::r2(m_flume)
#confint(m_flume, oldNames = FALSE)
r2_nakagawa(m_flume)
icc(m_flume)

#full_df$trt_lnppt <- interaction(full_df$Treatment, full_df$ln_ppt)
#loadR <- load_weight_time %>%
  #filter(!is.na(ro_event)) %>%
#    subset(random == 'R')
#m2_flume <- lmerTest::lmer(log(tss_sum+0.000359) ~ 
#                            Treatment*ln_ppt +
#                            Year*Treatment +
#                            
#                            #(1 | SiteID) + #removed due to singular fit
#                             #(1 | SiteID:Treatment) + 
#                            #(1|Year:Time) + #consider this and below with SiteID 
#                            (1|SiteID:Year:Time), 
#                          
#                          data = loadR)

#summary(m2_flume)
#anova(m2_flume)
#r2(m2_flume)
#icc(m2_flume)
#contr <- glht(m2_flume, linfct=mcp(trt_lnppt='Tukey'))
#summary(contr, test=adjusted("Shaffer"))

```

```{r design_step_model, dependson = "design_model"}
m_flume_step <- step(m_flume, reduce.random = FALSE, alpha.fixed = 0.1)
m_flume_model <- get_model(m_flume_step)
summary(m_flume_model)
anova(m_flume_model)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
```

#```{r design_step_model, dependson = "design_model"}
#emmip(m_flume, ln_ppt ~ Treatment | Year)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
#```


```{r design_model_contrasts, dependson = "design_model"}
trt_yr = emmeans(m_flume, pairwise ~ Treatment|Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(trt_yr)$contrasts
trt_yr$contrasts

trt = emmeans(m_flume, pairwise ~ Treatment, 
                    type = "response", 
                    lmer.df = "asymptotic")
confint(trt)

year = emmeans(m_flume, ~ Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(year)

##make this a function of what it is equal to (ln_ppt), could automate, but doing manual, ##uniroot()
trt_ppt = emmeans(m_flume, pairwise ~ Treatment|ln_ppt,
                    at=list(ln_ppt=c(-4,-3.5,-3.05,-3,-2)), 
                    #at=list(ln_ppt=c(-2.815)), 
                    type = "response",
                    lmer.df = "asymptotic")

contrast(trt_ppt, "pairwise")
confint(trt_ppt)$contrasts ## exp. the values

trt_ppt2 = emmeans(m_flume, "Treatment",
                   by='ln_ppt',
                    at=list(ln_ppt=c(-2.815, -3, -2.45, -3.62)), 
                    type = "response",
                    lmer.df = "asymptotic")
con = contrast(trt_ppt2, "pairwise", options= list(adjust = 'bonferroni'))
summary(con)
con
#emtrends(m_flume, pairwise~Treatment, var="ln_ppt")
#emmip(m_flume, Treatment~ln_ppt, cov.reduce = range)
#testInteractions(m_flume, residual=c("Treatment", "crop"))

#pred <- exp(predictInterval(m_flume, newdata = full_df, n.sims = 1000))

crop      = emmeans(m_flume, pairwise ~ Treatment|crop,   
                    type = "response",
                    lmer.df = "asymptotic")

confint(crop)$contrasts

```


```{r mean_tss_figure, dependson="load"}
# crop_em <- as.data.frame(crop)
#  
# k <- crop_em %>%
#   filter(contrast != "strips - control")
#  
# crop_em_plot <- k %>%
#    ggplot(aes(x=Treatment, y=response, fill=crop))+
#    geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
#    geom_errorbar(aes(ymin = (response-SE), ymax = (response+SE)),
#              width = 0.2,
#              linetype = "solid",
#              position = position_dodge(width = 0.5),
#              color="black", size=0.7) +
#    scale_fill_manual(values = c("corn" = "#FFCC33",
#                               "soybean" = "#339900"), 
#                     labels = c("Corn", "Soybean")) +
#    ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset)") +
#    xlab("Treatment") +
#    ylab("TSS Load \n(kg/ha)") +
#    theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
#          axis.title.x = element_text(size=18, face="bold"),
#          axis.title.y = element_text(size=18, face="bold"),
#          axis.text.x = element_text(size=18),
#          axis.text.y = element_text(size=18)) +
#   scale_x_discrete(labels= c("Control", "Prairie Strip")) #+
#   #scale_y_continuous(limits = c(0, 40))
#  
# crop_em_plot
# 
group <- full_df %>%
  group_by(Treatment) %>%
  summarize(n = n(),
            mean = mean(tss_sum),
            sd = sd(tss_sum),
            .groups = "drop") %>%
  mutate(se = sd / sqrt(n),
         lb = mean + qt(0.025, df = n-1)*se,
         ub = mean - qt(0.025, df = n-1)*se)

trt_plot <- group %>%
  ggplot(aes(x=Treatment, y=mean, fill=Treatment))+
  geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
  geom_errorbar(aes(ymin = (mean-se), ymax = (mean+se)),
                width = 0.3,
                linetype = "solid",
                position = position_dodge(width = 0.5),
                color="black", size=0.7) +
  scale_fill_manual(values = c("control" = "#C38820",
                              "strips" = "#176D9C")) +
  ggtitle("A. Full Dataset") +
  xlab("Treatment") +
  ylab("TSS Load (kg/ha)") +
 theme(plot.title = element_text(family="serif", size=28, face="bold", hjust=0.5),
       axis.title.x = element_text(family="serif", size=28, face="bold"),
       axis.title.y = element_text(family="serif", size=28, face="bold"),
       axis.text.x = element_text(family="serif", size=28),
       axis.text.y = element_text(family="serif", size=28)) +
  scale_x_discrete(labels= c("Prairie Strip", "Control"))+
  scale_y_continuous(limits = c(0, 40))

trt_plot
ggsave("E:/ISU/Project/SoilMove/data/statistics/flume_analysis/code/fig/flume_plot.png", trt_plot)
```

```{r mean_tss_figure, dependson="load"}
trtyr_em <- as.data.frame(trt_yr)
 
k <- trtyr_em %>%
  filter(contrast != "strips - control")
 
trtyr_em_plot <- k %>%
   ggplot(aes(x=Year, y=response, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (response-SE), ymax = (response+SE)),
             width = 0.2,
             linetype = "solid",
             position = position_dodge(width = 0.5),
             color="black", size=0.7) +
  scale_fill_manual(values = c("control" = "#C38820","strips" = "#176D9C")) +
   ggtitle("Comparison of Total Suspended Sediment (TSS) loads \n (Full Dataset)") +
   xlab("Year") +
   ylab("Ratio TSS Load \n(kg/ha)") +
   theme(plot.title = element_text(size=20, face="bold", hjust=0.5),
         axis.title.x = element_text(size=18, face="bold"),
         axis.title.y = element_text(size=18, face="bold"),
         axis.text.x = element_text(size=18),
         axis.text.y = element_text(size=18)) #+
  #scale_x_discrete(labels= c("Control", "Prairie Strip")) #+
  #scale_y_continuous(limits = c(0, 40))
 
trtyr_em_plot

trtyr_grp <- full_df %>% 
  group_by(Year, Treatment) %>% 
   summarize(n = n(), 
             mean = mean(tss_sum), 
             sd = sd(tss_sum),
             groups = "drop") %>%
   mutate(se = sd / sqrt(n),
          lb = mean + qt(0.025, df = n-1)*se,
          ub = mean - qt(0.025, df = n-1)*se)
 
 trt_plot <- trtyr_grp %>%
   ggplot(aes(x=Year, y=mean, fill=Treatment))+
   geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
   geom_errorbar(aes(ymin = (mean-se), ymax = (mean+se)),
                 width = 0.3,
                 linetype = "solid",
                 position = position_dodge(width = 0.5),
                 color="black", size=0.7) +
   scale_fill_manual(values = c("control" = "#C38820",
                               "strips" = "#176D9C")) +
   ggtitle("A. Full Dataset") +
   xlab("Treatment") +
   ylab("TSS Load (kg/ha)") +
  theme(plot.title = element_text(family="serif", size=20, face="bold", hjust=0.5),
        axis.title.x = element_text(family="serif", size=20, face="bold"),
        axis.title.y = element_text(family="serif", size=20, face="bold"),
        axis.text.x = element_text(family="serif", size=20),
        axis.text.y = element_text(family="serif", size=20)) +
   #scale_x_discrete(labels= c("Prairie Strip", "Control"))+
   scale_y_continuous(limits = c(0, 125))
 
 trt_plot
ggsave("E:/ISU/Project/SoilMove/data/statistics/flume_analysis/code/fig/flumeyr_plot.png", trt_plot)

```



# Check assumptions

There are two possible models:

- m_flume: full model design, design-based analysis
- m_flume_model: model design selected based on backward step selection

### Full model design

```{r design_model_residual_plots, dependson = "design_model"}
resid_panel(m_flume, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(m_flume)
```

### Selected model design
```{r design_step_residual_plots, dependson = "design_step_model"}
resid_panel(m_flume_model, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(m_flume_model)
```



```{r extract-contrast-ci, dependson="design_model"}
extract_contrast_ci <- function(d) {
  m <- lmerTest::lmer(update(m_flume, 
                          as.formula(paste0("~.+",d$variable))),
                    data = full_df) 
    
  # Obtain coefficient for new variable
  av <- anova(m)
  rowid <- which(rownames(av) == d$variable)
  
   
  em <- emmeans(m, 
                pairwise ~ Treatment, 
                type = "response", 
                lmer.df = "asymptotic")
  
  cbind(
    av[rowid,] %>% as.data.frame,
    confint(em)$contrasts %>% as.data.frame
  )
}
```


```{r treatment-effect-with-explanatory-variables, dependson=c("extract-contrast-ci")}
add <- data.frame(variable = c("ln_slp", "C_factor", "K_factor", "LSfactor", "Lfact_below", "LSfact_below", "Lfactor", "Sfactor", "ln_S", "ln_L", "ln_LS", "slope75", "ln_postL", "ln_postT"))

added_variable_analyses <- add %>% 
  group_by(variable) %>%
  do(extract_contrast_ci(.))
```


```{r print-added-variable-analysis, dependson="treatment-effect-with-explanatory-variables", results='asis'}
added_variable_analyses %>%
  xtable() %>%
  print(type="html", include.rownames=FALSE)
```


```{r caterpillar, eval = FALSE, echo = FALSE}

ggCaterpillar <- function(re, QQ=TRUE, likeDotplot=TRUE, reorder=TRUE) {
  require(ggplot2)
  f <- function(x) {
    pv   <- attr(x, "postVar")
    cols <- 1:(dim(pv)[1])
    se   <- unlist(lapply(cols, function(i) sqrt(pv[i, i, ])))
    if (reorder) {
      ord  <- unlist(lapply(x, order)) + rep((0:(ncol(x) - 1)) * nrow(x), each=nrow(x))
      pDf  <- data.frame(y=unlist(x)[ord],
                         ci=1.96*se[ord],
                         nQQ=rep(qnorm(ppoints(nrow(x))), ncol(x)),
                         ID=factor(rep(rownames(x), ncol(x))[ord], levels=rownames(x)[ord]),
                         ind=gl(ncol(x), nrow(x), labels=names(x)))
    } else {
      pDf  <- data.frame(y=unlist(x),
                         ci=1.96*se,
                         nQQ=rep(qnorm(ppoints(nrow(x))), ncol(x)),
                         ID=factor(rep(rownames(x), ncol(x)), levels=rownames(x)),
                         ind=gl(ncol(x), nrow(x), labels=names(x)))
    }

    if(QQ) {  ## normal QQ-plot
      p <- ggplot(pDf, aes(nQQ, y))
      p <- p + facet_wrap(~ ind, scales="free")
      p <- p + xlab("Standard normal quantiles") + ylab("Random effect quantiles")
    } else {  ## caterpillar dotplot
      p <- ggplot(pDf, aes(ID, y)) + coord_flip()
      if(likeDotplot) {  ## imitate dotplot() -> same scales for random effects
        p <- p + facet_wrap(~ ind)
      } else {           ## different scales for random effects
        p <- p + facet_grid(ind ~ ., scales="free_y")
      }
      p <- p + xlab("Levels") + ylab("Random effects")
    }

    p <- p + theme(legend.position="none")
    p <- p + geom_hline(yintercept=0)
    p <- p + geom_errorbar(aes(ymin=y-ci, ymax=y+ci), width=0, colour="black")
    p <- p + geom_point(aes(size=1.2), colour="blue") 
    return(p)
  }

  lapply(re, f)
}
```

```{r ranef, dependson = "design_model", eval = FALSE, echo = FALSE}
ranef(m_flume)
str(rr1 <- ranef(m_flume))

str(dd <- as.data.frame(rr1))
effects <- dd %>% 
  separate(grp, c('SiteID', 'Year', 'sample_event'))

dotplot(rr1, scales=list(x=list(relation='free')))[['SiteID:Year:sample_event']]
ggCaterpillar(ranef(m_flume,condVar=TRUE), QQ=TRUE, likeDotplot= TRUE, reorder=TRUE)[['SiteID:Year:sample_event']]

qqmath(rr1)

```


```{r ranef_plot, dependson = "ranef", eval = FALSE, echo = FALSE}
g <- ggplot(effects %>% filter(grpvar=='SiteID:Year:sample_event'), 
           aes(x = sample_event, y = condval)) + 
  #geom_line() +
  geom_point() +
  facet_grid(SiteID ~ Year) + 
  #scale_y_log10() + 
  labs(title ="Total load (log(kg/ha)) by sample event") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

```
