---
title: "Flume: Full Analysis"
subtitle: "(adapted from Jarad Niemi - Soilpad Analysis)"
author: "Jessica Nelson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("lme4")
library("lmerTest")
library("tidyverse"); theme_set(theme_bw())

library("emmeans")
library("ggResidpanel")
library("data.table")
library("stringr")


options(width = 120)

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```

## Read in data

```{r, child="sediment.Rmd"}
```


```{r load, dependson= "full_dataframe"}
load("full_df.RData")


```

## Exploratory analysis

## Site-year with sample event

```{r, dependson="load"}
site_year_rfevent <- full_df %>%
  select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_rfevent, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```

## Data visualization

## Number of samples

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r sample_event_counts, dependson="load"}
TSS_counts <- full_df %>%
  group_by(Year, SiteID, Treatment, crop, random) %>%
  distinct() %>%
  summarize(n = n(), .groups = "drop")
```

Plot the number of observations for each combination. 

```{r plot_count, dependson="sample_event_counts", out.width = "\\textwidth"}
g = ggplot(TSS_counts, aes(x = Treatment, y = crop, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  facet_grid(SiteID + random ~ Year) +
  scale_fill_gradient(low = "blue", high = "black") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g 
```

```{r, dependson="load"}
h <- ggplot(full_df, aes(x=ln_ppt, y=ln_tss_load, color=Treatment), inherit.aes = FALSE) +
  geom_point() + 
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
  ggtitle("Log-log relationship between TSS load and rainfall accumulation \n(full dataset)") +
  theme(plot.title = element_text(size=14, face="bold",hjust = 0.5))
  
h
#ggsave("fig/randReg_ppt_load.png", h, width = 12, height = 12)

```

```{r mean_data, dependson = "load"}
load_sum <- full_df %>%
  group_by(Year, SiteID, Treatment, crop) %>%
  summarize(sum_load             = sum(tss_load, na.rm = TRUE),
            log_load = mean(log(sum_load), na.rm = TRUE),
            n                   = n(),
            .groups = "drop")


ppt_sum <- flume %>%
  group_by(Year, SiteID) %>%
  summarize(sum_ppt             = sum(precipitation, na.rm = TRUE),
            #log_load = mean(log(sum_load), na.rm = TRUE),
            n                   = n(),
            .groups = "drop")

TSS_sum <- full_df %>%
  group_by(Year, SiteID, Treatment) %>%
  distinct() %>%
  summarize(tss_sum = sum(tss_load), .groups = "drop")

```

```{r wp_plot, dependson = "plot_data", eval = FALSE, echo = FALSE}
g <- ggplot(full_df, 
           aes(x = ln_ppt, y = ln_tss_load, group=Treatment, color=Treatment)) + 
  geom_line() +
  geom_point() +
  facet_grid(SiteID + random ~ Year) + 
  #scale_y_log10() + 
  labs(title ="Total load (log(kg/ha)) by rainfall accumulation per sample event (log(m))") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_plot.png", g)
```

```{r wp_per_day_plot, dependson = "plot_data"}
g <- ggplot(load_sum, 
           aes(x = Treatment, 
               y = sum_load)) + 
  geom_point() + 
  geom_line() +
  facet_grid(SiteID ~ Year, scales = "free_y") + 
  #scale_y_log10() + 
  labs(title ="Total load (kg/ha) per field season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_per_day_plot.png", g)
```

Average isn't realistic


```{r strip_boxplot, dependson = "create_soilpad", eval = FALSE, echo = FALSE}
g <- ggplot(flume, 
           aes(x = Year, y = precipitation, color=crop)) + 
  geom_boxplot() +
  geom_point() + 
  #geom_jitter() +
  facet_grid( ~ SiteID) + 
  #scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
g
```

#```{r, dependson="create_sediment"}
#pivot_sample %>% 
#  anova_test(ln_trt ~ ln_ppt*ln_ctl)
## purr https://stackoverflow.com/questions/50702152/compare-models-via-anova-with-purrr-or-dplyr
## anova() and may need an linear model built up.
#```


# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss

## Confirmatory, design-based analysis

### Treatment effect

```{r design_model, dependson = "load"}
m_flume <- lmerTest::lmer(log(tss_load+0.00053) ~ 
                             Treatment*ln_ppt +
                            
                            Year*Treatment +
                            
                            #(1 | SiteID) + #removed due to singular fit
                             (1 | SiteID:Treatment) + 
                            #(1|Year:sample_event) + #consider this and below with SiteID 
                            (1|SiteID:Year:sample_event), 
                          
                          data = full_df)

summary(m_flume)

```

```{r design_step_model, dependson = "design_model"}
m_flume_step <- step(m_flume, reduce.random = FALSE, alpha.fixed = 0.1)
m_flume_model <- get_model(m_flume_step)
summary(m_flume_model)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
```

#```{r design_step_model, dependson = "design_model"}
#emmip(m_flume, ln_ppt ~ Treatment | Year)

##https://campus.datacamp.com/courses/hierarchical-and-mixed-effects-models-in-r/linear-mixed-effect-models?ex=7
#```


```{r design_model_contrasts, dependson = "design_model"}
trt_yr = emmeans(m_flume, pairwise ~ Treatment|Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(trt_yr)$contrasts

trt = emmeans(m_flume, pairwise ~ Treatment, 
                    type = "response", 
                    lmer.df = "asymptotic")
confint(trt)

year = emmeans(m_flume, ~ Year, 
                    type = "response",
                    lmer.df = "asymptotic") 
confint(year)

trt_ppt = emmeans(m_flume, pairwise ~ Treatment|ln_ppt,
                    at=list(ln_ppt=c(-4,-3.5,-3,-2)), 
                    type = "response",
                    lmer.df = "asymptotic")

confint(trt_ppt)$contrasts ## exp. the values

#crop      = emmeans(m_flume, pairwise ~ Treatment|crop,   
#                    type = "response",
#                    lmer.df = "asymptotic")

#confint(crop)$contrasts

```

```{r diversion_figure, dependson="diversion_contrasts"}
trt <- as.data.frame(trt)

k <- trt %>%
  filter(contrast != "strips - control")

trt_plot <- k %>%
  ggplot(aes(x=Treatment, y=response, fill=Treatment))+
  geom_bar(width = 0.5, position = position_dodge(), stat="summary") +
  geom_errorbar(aes(ymin = (response-SE), ymax = (response+SE)),
            width = 0.2,
            linetype = "solid",
            position = position_dodge(width = 0.5),
            color="black", size=0.7) +
  scale_fill_manual(values = c("control" = "#176D9C",
                               "strips" = "#C38820")) +
  ggtitle("Comparison of Total Suspended Sediment (TSS) loads") +
  xlab("Treatment") +
  ylab("TSS Load \n(Mg/ha))") +
  theme(plot.title = element_text(size=28, face="bold", hjust=0.5),
        axis.title.x = element_text(size=20, face="bold"),
        axis.title.y = element_text(size=20, face="bold"),
        axis.text.x = element_text(size=20),
        axis.text.y = element_text(size=20)) +
  scale_x_discrete(labels= c("Control", "Prairie Strip"))

trt_plot
```

# Check assumptions

There are two possible models:

- m_flume: full model design, design-based analysis
- m_flume_model: model design selected based on backward step selection

### Full model design

```{r design_model_residual_plots, dependson = "design_model"}
resid_panel(m_flume, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(m_flume)
```

### Selected model design
```{r design_step_residual_plots, dependson = "design_step_model"}
resid_panel(m_flume_model, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(m_flume_model)
```