---
title: "analysis_nonrandom"
author: "Jnelly"
date: "3/16/2022"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("lme4")
library("lmerTest")
library("tidyverse"); theme_set(theme_bw())
library("emmeans")
library("ggResidpanel")
library("data.table")
library("stringr")

options(width = 120)

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```

## Read in data

```{r, child="sediment.Rmd"}
```

```{r, diversion_coding, dependson = "create_soilpad"}
flumeNR <- TSS_events %>%
  #filter(!is.na(ro_event)) %>%
    subset(random == 'NR')

```

## Exploratory analysis

## Site-year with rainfall event

```{r, dependson="create_soilpad"}
site_year_rfeventNR <- flumeNR %>%
  select(SiteID, Year, rf_event) %>%
  unique()

ggplot(site_year_rfeventNR, aes(Year, SiteID, fill=rf_event)) +
  geom_tile()
```

## Number of samples

Calculate the number of observations for each treatment-position-year-site-time
combination. 

```{r soilpad_counts, dependson="create_soilpad"}
TSS_countsNR <- flumeNR %>%
  group_by(Year, SiteID, rf_event) %>%
  distinct() %>%
  summarize(n = n(), .groups = "drop")
```

Plot the number of observations for each combination. 

```{r plot_count, dependson="soilpad_counts", out.width = "\\textwidth"}
g <- ggplot(TSS_countsNR, aes(x = Year, y= SiteID, fill = n)) +
  geom_tile() +
  geom_text(aes(label = n), color = "white") +
  scale_fill_gradient(low = "blue", high = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

g 
##ggsave("fig/soilpad_counts_no_diversion.png", g, width = 12, height = 12)
```
# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss

```{r soilpad_counts, dependson="create_soilpad"}
flumeR_sum <- flumeNR %>%
  group_by(Treatment, Year, SiteID, ro_event, tss_sum) %>%
  summarize(ppt_sum = sum(precipitation)) %>%
  distinct()

```


```{r design_model, dependson = "create_soilpad"}

mNR_flume <- lmerTest::lmer(log(tss_sum+0.000198) ~ 
                             (1 | SiteID) +
                             #(1 | SiteID:Treatment) + 
                             (1 | SiteID:ppt_sum) + 
                             Treatment*ppt_sum,
                           data = flumeNR_sum)

mNR_flume_step <- step(mNR_flume, 
                    reduce.random = FALSE,
                    alpha.fixed = 0.1)

mNR_flume_step

mNR_flume_model <- get_model(mNR_flume_step)
# design_model <- design_model_full # override backward selection

summary(mNR_flume_model)
```