---
title: "Flume: Full Analysis (March-November)"
subtitle: "(adapted from Jarad Niemi - Soilpad Analysis)"
author: "Jessica Nelson"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      fig.width = 12,
                      fig.height = 12)

library("tidyverse"); theme_set(theme_bw())
library(cowplot) #for manuscript ready figures
library(lme4) #for lmer & glmer models
library(sjPlot) #for plotting lmer and glmer mods
library(sjmisc) 
library(effects)
library(sjstats)

options(width = 120, scipen = 999)

dir.create("fig", showWarnings = FALSE)
```

```{r}
sessionInfo()
```


```{r load, dependson= "full_dataframe"}
load("E:/ISU/Project/SoilMove/data/statistics/flume_analysis/code/full_df.RData")

rain_event <- read.csv('E:/ISU/Project/SoilMove/data/statistics/flume_analysis/data/tidy/rain_event12.csv', header=TRUE)

rain_event <- rain_event %>% filter(!duplicated(cbind(Year, SiteID, eventStart)))

runoff_event <- read.csv('E:/ISU/Project/SoilMove/data/statistics/flume_analysis/data/tidy/runoff_event12.csv', header=TRUE)

```

## Exploratory analysis

## Site-year with sample event

```{r, dependson="load"}
site_year_rfevent <- full_df %>%
  select(SiteID, Year, sample_event) %>%
  unique()

ggplot(site_year_rfevent, aes(Year, SiteID, fill=sample_event)) +
  geom_tile()
```

```{r mean_data, dependson = "load"}
ppt_sum <- rain_event %>%
  filter(!duplicated(cbind(Year, SiteID, eventStart))) %>%
  group_by(Year, SiteID) %>%
  summarize(sum_pptmm = sum(precipitation, na.rm = TRUE)*1000,
            time_min = sum(rain_time),
            n = n(),
            .groups = "drop")

runoff_sum <- runoff_event %>%
  filter(!duplicated(cbind(Year, SiteID, eventStart))) %>%
  group_by(Year, SiteID, Treatment) %>%
  distinct() %>%
  summarize(sum_flow = sum(flow),
            time_min = sum(flow_time),
            n = n(),
            .groups = "drop")

runoff_mean <- runoff_event %>%
  filter(!duplicated(cbind(Year, SiteID, eventStart))) %>%
  group_by(Year, SiteID, Treatment) %>%
  distinct() %>%
  summarize(sum_flow = sum(flow),
            time_min = sum(flow_time),
            n = n(),
            .groups = "drop")

```

```{r wp_plot, dependson = "plot_data", eval = FALSE, echo = FALSE}
g <- ggplot(full_df, 
           aes(x = log(ppt_sum), y = log(tss_load+0.0016488035), group=Treatment, color=Treatment)) + 
  geom_line() +
  geom_point() +
  facet_grid(SiteID + random ~ Year) + 
  #scale_y_log10() + 
  labs(title ="Total load (log(kg/ha)) by rainfall accumulation per sample event (log(m))") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_plot.png", g)
```



```{r wp_per_day_plot, dependson = "plot_data"}
g <- ggplot(load_sum, 
           aes(x = Treatment, 
               y = sum_load)) + 
  geom_point() + 
  geom_line() +
  facet_grid(SiteID ~ Year, scales = "free_y") + 
  #scale_y_log10() + 
  labs(title ="Total load (kg/ha) per field season") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
g

#ggsave("fig/wp_per_day_plot.png", g)
```

Average isn't realistic


```{r strip_boxplot, dependson = "create_soilpad", eval = FALSE, echo = FALSE}
g <- ggplot(flume, 
           aes(x = Year, y = precipitation, color=crop)) + 
  geom_boxplot() +
  geom_point() + 
  #geom_jitter() +
  facet_grid( ~ SiteID) + 
  #scale_y_log10() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))
g
```

#```{r, dependson="create_sediment"}
#pivot_sample %>% 
#  anova_test(ln_trt ~ ln_ppt*ln_ctl)
## purr https://stackoverflow.com/questions/50702152/compare-models-via-anova-with-purrr-or-dplyr
## anova() and may need an linear model built up.
#```


# Main Analyses

There are three main analyses of interest:

- confirmatory, design-based analysis
- exploratory, covariate analysis
- relationship of sediment flow to sediment loss

## Confirmatory, design-based analysis

### Treatment effect

```{r design_model, dependson = "load"}
mf_paper <- lme4::lmer(log(tss_load+0.0016) ~ 
                         Treatment*ln_ppt +
                         #Treatment*crop +
                         Year*Treatment +
                         #(1|SiteID) + #removed due to singular fit
                         (1|SiteID:Treatment) + 
                         (1|Year:sample_event) + #consider this and below with SiteID
                         (1|SiteID:Year:sample_event), 
                       REML=FALSE,
                       data = full_df)

summary(mf_paper)

```

# Check assumptions

There are two possible models:

- m_flume: full model design, design-based analysis
- m_flume_model: model design selected based on backward step selection

### Full model design

```{r design_model_residual_plots, dependson = "design_model"}
resid_panel(mf_paper, 
            plots = c("resid","index","yvp","qq"),
            smoother = TRUE, qqbands = TRUE)

resid_xpanel(mf_paper)
```

## Plot of the effect sizes
```{r design_model_contrasts, dependson = "design_model"}
sjPlot::plot_model(mf_paper)

```

```{r diversion_figure, dependson="diversion_contrasts"}
sjPlot::plot_model(mf_paper,
                   show.values=TRUE, show.p=TRUE,
                   title='Sediment transport predictor variables \n (RandomDataset)')
```

```{r diversion_figure, dependson="diversion_contrasts"}
sjPlot::tab_model(mf_paper,
                  show.re.var=TRUE,
                  pred.labels = c("Intercept", "Control Treatment", "log(rainfall (mm))", "2017",
                                  "2018", "2019", "2020", "2021", "Control*log(rainfall)", "Control*2017",
                                  "Control*2018", "Control*2019",
                                  "Control*2020", "Control*2021"),
                  dv.labels = "Full Dataset")

```

```{r diversion_figure, dependson="diversion_contrasts"}
effects_trt <- effects::effect(term= "Treatment", mod= mf_paper)
summary(effects_trt) #output of what the values are
```

```{r diversion_figure, dependson="diversion_contrasts"}
effects_ppt <- effects::effect(term= "ln_ppt", mod= mf_paper)
summary(effects_trt) #output of what the values are

x_ppt <- as.data.frame(effects_ppt)

```


```{r diversion_figure, dependson="diversion_contrasts"}
urchin_plot <- ggplot() + 
  #2
geom_point(data=full_df, aes(ln_ppt, log(tss_load))) + 
  #3
  geom_point(data=x_ppt, aes(x=ln_ppt, y=fit), color="blue") +
  #4
  geom_line(data=x_ppt, aes(x=ln_ppt, y=fit), color="blue") +
  #5
  #geom_ribbon(data= x_ppt, aes(x=ln_ppt, ymin=lower, ymax=upper), alpha= 0.3, fill="blue") +
  #6
  labs(x="Urchins (centered & scaled)", y="Coral Cover")

urchin_plot

h <- ggplot(data = full_df, aes(y = log(tss_load), x = ln_ppt, fill = Treatment, color = Treatment)) + 
    #geom_line(aes(group = Treatment), col = "black") +
    geom_point(size = 3, position = position_dodge(width = 0.7), show.legend = FALSE) +
  geom_smooth(method=lm, se=FALSE, fullrange=TRUE) +
    facet_wrap(~ SiteID, nrow = 7) +
    labs(x = "Treatment", y = "log phenolics concentration") +
    theme(aspect.ratio = 2)+
    theme_classic()

h

visreg(mf_paper, xvar = "ln_ppt", by = "Treatment", overlay = TRUE,
      ylab="Log phenolics concentration", data = x)

```